---
title: "dottable homework"
author: "Gabriela C. Lebron & Alasya Zeweldi"
date: "2025-03-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(data.table)
```

# Go to GitHub and download the data nycdata.csv and then answer the following questions.

```{r}
library(RCurl)
x <- getURL("https://raw.githubusercontent.com/semiyarih/DATA613/refs/heads/main/dottable/nycdata.csv")
flights <- read.csv(text = x)
flights.data <- data.table(flights)
head(flights, 6)
```


## 1.Use and show data.table code to select the variables year, month, day, and hour from the imported flights data

```{r}
flights.data[, .(year, month, day, hour)]
```


## 2. Use and show data.table code to produce a table that shows a carrier of DL, an origin of JFK and a destination of SEA

```{r}
flights.data[carrier == "DL" & origin == "JFK" & dest == "SEA"]
```


## 3. Use and show data.table code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.

```{r}
flights.data[carrier == "UA" & month == 3 & air_time < 330]
```

## 4. Use and show tidyverse code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.

```{r}
flights %>% 
  filter(carrier == "UA" & month == 3 & air_time < 330) %>% 
  head(6)
```

## 5. Use the data.table method to add a variable called speed that is the average air speed of the plane in miles per hour.

```{r}
flights.data[,speed := distance/hour]
head(flights.data)
```

## 6. Use the tidyverse method to add a variable called speed that is the average air speed of the plane in miles per hour.

```{r}
flights %>% 
  mutate(speed = distance/hour) %>% 
  head(6)
```

## 7. Show and use coding to change the carrier abbreviation of UA to UniitedAir,

### a. data.table method

```{r}
flights.data[carrier == "UA", carrier := "UniitedAir"]
unique(flights.data$carrier)
```

### b. tidyverse method (Use a sequence of dplyr commands so that you can see the change in your table)

```{r}
flights %>%
  mutate(carrier = ifelse(carrier == "UA", "UniitedAir", carrier)) %>% 
  filter(carrier == "UniitedAir") %>% 
  head(6)

```

